Designing Types for R, Empirically

### 2020-05-15 paper/main.tex
minor

\begin{lstlisting}
binom.profile <- function(x, n, ...) {
  xn <- cbind(x = x, n = n)
  ok <- !is.na(xn[, 1]) & !is.na(xn[, 2])
  x <- xn[ok, "x"]
  n <- xn[ok, "n"]
  ...
\end{lstlisting}

@@

\begin{lstlisting}
hankel.matrix <- function( n, x ) { 
### arguments
### n = a positive integer value for the order of the Hankel matrix
### x = an order 2 * n + 1 vector of numeric values
    ...
    if ( n != trunc( n ) )  stop( "argument n ix not an integer" )
    if ( !is.vector( x ) )  stop( "argument x is not a vector" )
    m <- length( x )
    if ( m < n )  stop( "length of argument x is less than n" )
    ...
\end{lstlisting}

@@

\begin{lstlisting}
rowWeightedMeans <- function(x, w = NULL, rows = NULL, cols = NULL,
                             na.rm = FALSE, ...) {
  # - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
  # Validate arguments
  # - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
  # Argument 'x':
  if (!is.matrix(x)) {
    .Defunct(msg = sprintf("Argument 'x' is of class %s, but should be a matrix. 
    The use of a %s is not supported, the correctness of the result is not guaranteed. 
    Please update your code accordingly.", sQuote(class(x)[1]), sQuote(class(x)[1])))
  }
  ...
\end{lstlisting}

@@

\begin{lstlisting}
rowWeightedMeans <- function(x, w = NULL, rows = NULL, cols = NULL,
                             na.rm = FALSE, ...) {
  ...
  n <- ncol(x)
  if (length(w) != n) {
    stop("The length of argument 'w' is does not match the number of column in 'x': ", 
    length(w), " != ", n)
  }
  ...
\end{lstlisting}

@@

\paragraph{Structs.}

@@

\begin{lstlisting}
agricolae::cv.model <- function(x) {
  suma2 <- sum(x$residual^2)
  gl <- x$df.residual
  promedio <- mean(x$fitted.values)
  return(sqrt(suma2/gl)*100/promedio)
}

data(sweetpotato)
model<-aov(yield~virus, data=sweetpotato)
cv.model(model)
\end{lstlisting}

@@

\begin{lstlisting}
data(cars)

# tidyverse: Filter rows where speed is > 20. 
fast_cars_tidy <- filter(cars, speed > 20)

# base R: Get all rows where the "speed" column has a value > 20.
fast_cars_base <- cars[cars[, "speed"] > 20, ] # base example

\end{lstlisting}

@@

We breakdown the failed assertions by type in
Table~\ref{tbl:fail-breakdown-0}.  By far the most commonly failed assertion
is one asserting that a value with ${\bf dbl}[]$ type has type ${\bf
  class}\langle matrix \rangle$.  This was a deliberate design decision, as
coercion from vectors to matrices is ad hoc at best, and surely not a
practice codified in the language.  In a similar vein, another popular
failing assertion is checking if a ${\bf dbl}[]$ has type ${\bf int}[]$,
another case of commonly performed coercion.  We did not include these types
of coercions in our type annotation framework as programmers cannot rely on
them, and it is not always the case that the coercions are safe to perform.
=======

>>>>

\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
  binom.profile <- function(x, n, ...) {
   xn <- cbind(x = x, n = n)
   ok <- !is.na(xn[, 1]) & !is.na(xn[, 2])
   x <- xn[ok, "x"]
   n <- xn[ok, "n"]
\end{lstlisting}

@@

\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
  hankel.matrix <- function( n, x ) { 
  ### n = a positive integer value for the order of the Hankel matrix
  ### x = an order 2 * n + 1 vector of numeric values
    ...
    if ( n != trunc( n ) )  stop( "argument n ix not an integer" )
    if ( !is.vector( x ) )  stop( "argument x is not a vector" )
    m <- length( x )
    if ( m < n )  stop( "length of argument x is less than n" )
\end{lstlisting}

@@

\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
rowWeightedMeans <- function(x, w = NULL, rows = NULL, cols = NULL,
                             na.rm = FALSE, ...) {
  if (!is.matrix(x)) 
    .Defunct(msg = sprintf("Argument 'x' should be a matrix.)
\end{lstlisting}

@@

\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
  rowWeightedMeans <- function(x, w = NULL, rows = NULL, cols = NULL, na.rm = FALSE) {
    n <- ncol(x)
    if (length(w) != n) 
       stop("The length of 'w' does not match columns in 'x': ")
\end{lstlisting}

@@

\paragraph{Structs}

@@

\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
  agricolae::cv.model <- function(x) {
    suma2 <- sum(x$residual^2)
    gl <- x$df.residual
    promedio <- mean(x$fitted.values)
    return(sqrt(suma2/gl)*100/promedio)
  }
  data(sweetpotato)
  model<-aov(yield~virus, data=sweetpotato)
  cv.model(model)
\end{lstlisting}

@@

\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
data(cars)
# tidyverse: Filter rows where speed is > 20. 
fast_cars_tidy <- filter(cars, speed > 20)
# base R: Get all rows where the "speed" column has a value > 20.
fast_cars_base <- cars[cars[, "speed"] > 20, ] # base example
\end{lstlisting}

@@

We breakdown the failed assertions by type in
Table~\ref{tbl:fail-breakdown-0}.  By far the most commonly failed assertion
is one asserting that a value with ${\bf dbl}[]$ type has type ${\bf
  class}\langle matrix \rangle$.  This was a deliberate design decision, as
coercion from vectors to matrices is ad hoc at best, and surely not a
practice codified in the language.  In a similar vein, another popular
failing assertion is checking if a ${\bf dbl}[]$ has type ${\bf int}[]$,
another case of commonly performed coercion.  We did not include these types
of coercions in our type annotation framework as programmers cannot rely on
them, and it is not always the case that the coercions are safe to perform.
